
---
title: "GSEA&KEGG"
author: "Aykhan Yusifov"
date: "7/30/2020"
output:
  html_document: null
  pdf_document: default
---

# Load Required R Packages


```{r}
#Install and load required packages
#BiocManager::install("clusterProfiler", version = "3.8")
#BiocManager::install("pathview")
#BiocManager::install("enrichplot")
library(clusterProfiler)
library(enrichplot)
# I use ggplot2 -easy tool for adding details
library(ggplot2)
```



```{r}

# Set my organism
# Please check from the list and find the organism you need(I am not sure what model you are using)
organism = "org.Mm.eg.db"
#BiocManager::install(organism, character.only = TRUE)
library(organism, character.only = TRUE)

```



```{r}

# loading the dataset
df = read.csv("YourData.csv", header=TRUE)

# I want the log fold change 
original_gene_list <- df$logFC

# name the vector
names(original_gene_list) <- df$Ensembl

# omit any NA values 
gene_list<-na.omit(original_gene_list)

# sort the list in decreasing order (required for clusterProfiler)
gene_list = sort(gene_list, decreasing = TRUE)


```


#Gene set Enrichment
# params -keyType This is the source of the annotation (gene ids). The options vary for each annotation. In the example of org.Mm.eg.db, the options are:“ACCNUM” “ALIAS” “ENSEMBL” “ENSEMBLPROT” “ENSEMBLTRANS” “ENTREZID” “ENZYME” “EVIDENCE” “EVIDENCEALL” “FLYBASE” “FLYBASECG” “FLYBASEPROT”“GENENAME” “GO” “GOALL” “MAP” “ONTOLOGY” “ONTOLOGYALL” “PATH” “PMID” “REFSEQ” “SYMBOL” “UNIGENE” “UNIPROT”. Check which options are available with the keytypes command, for example keytypes(org.Mm.eg.db).
 
#ont one of “BP”, “MF”, “CC” or “ALL”
#nPerm the higher the number of permutations you set, the more accurate your result will, but the longer the analysis will take.
#minGSSize minimum number of genes in set (gene sets with lower than this many genes in your dataset will be ignored).
#maxGSSize maximum number of genes in set (gene sets with greater than this many genes in your dataset will be ignored).
#pvalueCutoff pvalue Cutoff.
#pAdjustMethod one of “holm”, “hochberg”, “hommel”, “bonferroni”, “BH”, “BY”, “fdr”, “none”

```{r}
#Here I just use BP for example, for checking others, erase BP add CC or any other you want
gse <- gseGO(geneList=gene_list, 
             ont ="BP",
             keyType = "ENSEMBL", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = organism, 
             pAdjustMethod = "none")

```


#Dotplot


```{r, fig.width = 12}
require(DOSE)
dotplot(gse, showCategory=10, split=".sign") + facet_grid(.~.sign) + xlim(0, 1)


```


#KEGG Gene Set Enrichment Analysis

#For KEGG pathway enrichment using the gseKEGG() function, we need to convert id types. I use the bitr function for this (included in clusterProfiler). 

#bitr function has to be one of the available options from keyTypes(org.Mm.eg.db) and must map to one of ‘kegg’, ‘ncbi-geneid’, ‘ncib-proteinid’ or ‘uniprot’ because gseKEGG() only accepts one of these 4 options as it’s keytype parameter. In the case of org.Dm.eg.db, none of those 4 types are available, but ‘ENTREZID’ are the same as ncbi-geneid for org.Mm.eg.db so we use this for toType.

#As our intial input, we use original_gene_list which we created above.

#Prepare Input

# Convert gene IDs for gseKEGG function, it can get tricky here as your data set gene ids might not be compatible, if you have questions on that, let me know.

# We will lose some genes here because not all IDs will be converted

```{r}
ids<-bitr(names(original_gene_list), fromType = "ENSEMBL", toType = "ENTREZID", OrgDb=organism)
# Remove duplicate IDS (here I use "ENSEMBL", but it should be whatever was selected as keyType)
dedup_ids = ids[!duplicated(ids[c("ENSEMBL")]),]
```

# Here you can see ENsembl to entrezid correspondance
```{r}
dedup_ids
```
# To merge and make one list I am giving the same col name

```{r}
colnames(dedup_ids) <- c("Ensembl", "ENTREZID")
```


#Create a new dataframe df2 which has only the genes which were successfully mapped using the bitr function above and the column for entrezid

```{r}
df2 = merge (df, dedup_ids, by=c("Ensembl", "Ensembl"))

df2
```


#Create a vector of the gene unuiverse

```{r}
kegg_gene_list <- df2$logFC

```

#Name vector with ENTREZ ids
```{r}
names(kegg_gene_list) <- df2$ENTREZID
```


#omit any NA values 
```{r}
kegg_gene_list<-na.omit(kegg_gene_list)
```


#sort the list in decreasing order (required for clusterProfiler)
```{r}
kegg_gene_list = sort(kegg_gene_list, decreasing = TRUE)
```


#Create gseKEGG object
#KEGG Organism Code: The full list is here: https://www.genome.jp/kegg/catalog/org_list.html (need the 3 letter code). I define this as kegg_organism first, because it is used again below when making the pathview plots.

```{r}
kegg_organism = "mmu"

kk2 <- gseKEGG(geneList     = kegg_gene_list,
               organism     = kegg_organism,
               nPerm        = 10000,
               minGSSize    = 3,
               maxGSSize    = 800,
               pvalueCutoff = 0.05,
               pAdjustMethod = "none",
               keyType       = "ncbi-geneid")


```



#DotPlot


```{r, fig.width = 10}

dotplot(kk2, showCategory = 10, title = "Enriched Pathways" , split=".sign") + facet_grid(.~.sign) + scale_x_continuous(limits=c(0,1), labels = scales::number_format(accuracy = 0.2))


```

